---
title: "Multiomics Project"
format: pdf
editor: visual
---

```{r, echo=FALSE}
library(HDF5Array)
library(rhdf5)
```

## GSE71669

Condition: Invasive/Non-invasive Bladder Cancer

Number of samples: 33

Experiments: 3 experiments

```         
[1] mrna: HDF5Matrix with 19869 rows and 33 columns
[2] cpg: HDF5Matrix with 27237 rows and 33 columns
[3] cc: HDF5Matrix with 45 rows and 33 columns
```

Columns in .RDS:

-   sample name

-   response \[0/1\]

```{r}
# Load the MAE object from your .rds file
mae_GSE71669 <- readRDS("~/GSE71669/mae_data/mae.rds")
mae <- mae_GSE71669

# # Check the class of the object
# class(mae)
# 
# # Print the summary of the MAE object
# summary(mae)
# 
# # List all assays (experiments) contained in the MAE object
# experiments(mae)
# 
# # Get a quick overview of the metadata associated with samples (colData)
# colData(mae)
# metadata(mae)

# Open the HDF5 file
h5_file <- HDF5Array::H5File("~/GSE71669/mae_data/experiments.h5")
h5_file

# Load the HDF5 file containing assay data (if it's separate from the RDS)
h5_contents <- h5ls(h5_file)

# get assay data from mrna dataset 
assay_data <- h5mread(h5_file, "assay001")

h5closeAll()

# Read the dataset with gene and sample dimnames
gene_dimnames_GSE <- h5read("~/GSE71669/mae_data/experiments.h5", "/.assay001_dimnames/1")
sample_dimnames <- h5read("~/GSE71669/mae_data/experiments.h5", "/.assay001_dimnames/2")

# convert assay data into a dataframe with genes as columns and sample names as rows
class(assay_data)
GSE_assay_df <- as.data.frame(t(assay_data))
rownames(GSE_assay_df) <- sample_dimnames
colnames(GSE_assay_df) <- gene_dimnames_GSE
```

## ROSMAP

Condition: –- unknown

Number of samples: 351

Experiments: 3 experiments

```         
[1] epigenomics: HDF5Matrix with 200 rows and 351 columns
[2] genomics: HDF5Matrix with 200 rows and 351 columns
[3] transcriptomics: HDF5Matrix with 200 rows and 351 columns
```

Columns in .RDS:

-   sample name

-   response \[no/yes\]

```{r}
# Load the MAE object from your .rds file
mae_rosmap <- readRDS("~/rosmap/mae_data/mae.rds")
mae <- mae_rosmap

# # Check the class of the object
# class(mae)
# 
# # Print the summary of the MAE object
# summary(mae)
# 
# # List all assays (experiments) contained in the MAE object
# experiments(mae)
# 
# # Get a quick overview of the metadata associated with samples (colData)
# colData(mae)
# metadata(mae)

# Open the HDF5 file
h5_file <- HDF5Array::H5File("~/rosmap/mae_data/experiments.h5")
h5_file

# Load the HDF5 file containing assay data (if it's separate from the RDS)
h5_contents <- h5ls(h5_file)

# get assay data from mrna dataset 
assay_data <- h5mread(h5_file, "assay003")

h5closeAll()

# Read the dataset with gene and sample dimnames
gene_dimnames_rosmap <- h5read("~/rosmap/mae_data/experiments.h5", "/.assay003_dimnames/1")
gene_dimnames_rosmap_2 <- h5read("~/rosmap/mae_data/experiments.h5", "/.assay002_dimnames/1")
gene_dimnames_rosmap_3 <- h5read("~/rosmap/mae_data/experiments.h5", "/.assay001_dimnames/1")
sample_dimnames <- h5read("~/rosmap/mae_data/experiments.h5", "/.assay003_dimnames/2")

# convert assay data into a dataframe with genes as columns and sample names as rows
class(assay_data)
assay_df <- as.data.frame(t(assay_data))
rownames(assay_df) <- sample_dimnames
colnames(assay_df) <- gene_dimnames_rosmap
```

## TCGA-BLCA

Condition: Bladder Urothelial Carcinoma

Number of Samples: 336

Experiments: 4 experiments

```         
[1] Methylation_Gene_level_HM450K: HDF5Matrix with 20109 rows and 336 columns
[2] miRNA_Gene_level: HDF5Matrix with 825 rows and 336 columns
[3] RPPA_Gene_Level: HDF5Matrix with 155 rows and 336 columns
[4] RNAseq_HiSeq_Gene_level: HDF5Matrix with 20047 rows and 336 
```

Columns:

-   Age \[int\]

-   Sex \[0/1\]

-   Response \[0/1\]

```{r}
# Load the MAE object from your .rds file
mae_tcga_blca <- readRDS("~/tcga-blca/tcga-blca_mae_data/mae.rds")
mae <- mae_tcga_blca

# # Check the class of the object
# class(mae)
# 
# # Print the summary of the MAE object
# summary(mae)
# 
# # List all assays (experiments) contained in the MAE object
# experiments(mae)
# 
# # Get a quick overview of the metadata associated with samples (colData)
# colData(mae)
# metadata(mae)

# Open the HDF5 file
h5_file <- HDF5Array::H5File("~/tcga-blca/tcga-blca_mae_data/experiments.h5")
h5_file

# Load the HDF5 file containing assay data (if it's separate from the RDS)
h5_contents <- h5ls(h5_file)

# get assay data from mrna dataset 
assay_data <- h5mread(h5_file, "assay004")

h5closeAll()
# Read the dataset with gene and sample dimnames
gene_dimnames_blca <- h5read("~/tcga-blca/tcga-blca_mae_data/experiments.h5", "/.assay004_dimnames/1")
sample_dimnames <- h5read("~/tcga-blca/tcga-blca_mae_data/experiments.h5", "/.assay004_dimnames/2")

# convert assay data into a dataframe with genes as columns and sample names as rows
class(assay_data)
blca_assay_df <- as.data.frame(t(assay_data))
rownames(blca_assay_df) <- sample_dimnames
colnames(blca_assay_df) <- gene_dimnames_blca
```

## TCGA-BRCA

Condition: Breast invasive carcinoma

Number of Samples: 109

Experiments: 4 experiments

```         
[1] Methylation_Gene_level_HM450K: HDF5Matrix with 20106 rows and 109 columns
[2] miRNA_GA_Gene_level: HDF5Matrix with 754 rows and 109 columns
[3] RPPA_Gene_level: HDF5Matrix with 175 rows and 109 columns
[4] RNAseq_HiSeq_Gene_level: HDF5Matrix with 20155 rows and 109 co
```

Columns:

-   Age \[int\]

-   Sex \[0\]

-   Response \[0/1\]

```{r}
# Load the MAE object from your .rds file
mae_tcga_brca <- readRDS("~/tcga-brca/tcga-brca_mae_data/mae.rds")
mae <- mae_tcga_brca

# # Check the class of the object
# class(mae)
# 
# # Print the summary of the MAE object
# summary(mae)
# 
# # List all assays (experiments) contained in the MAE object
# experiments(mae)
# 
# # Get a quick overview of the metadata associated with samples (colData)
# colData(mae)
# metadata(mae)

# Open the HDF5 file
h5_file <- HDF5Array::H5File("~/tcga-brca/tcga-brca_mae_data/experiments.h5")
h5_file

# Load the HDF5 file containing assay data (if it's separate from the RDS)
h5_contents <- h5ls(h5_file)

# get assay data from mrna dataset 
assay_data <- h5mread(h5_file, "assay004")

h5closeAll()
# Read the dataset with gene and sample dimnames
gene_dimnames_brca <- h5read("~/tcga-brca/tcga-brca_mae_data/experiments.h5", "/.assay004_dimnames/1")
sample_dimnames<- h5read("~/tcga-brca/tcga-brca_mae_data/experiments.h5", "/.assay004_dimnames/2")

# convert assay data into a dataframe with genes as columns and sample names as rows
class(assay_data)
brca_assay_df <- as.data.frame(t(assay_data))
rownames(brca_assay_df) <- sample_dimnames
colnames(brca_assay_df) <- gene_dimnames_brca
```

## TCGA-KIPAN

Condition: Pan-kidney cohort (Kidney Chromophobe + Kidney renal clear cell carcinoma + Kidney renal papillary cell carcinoma)

Number of Samples: 123

Experiments: 4 experiments

```         
 [1] Methylation_Gene_level_HM450K: HDF5Matrix with 20116 rows and 123 columns
 [2] miRNA_GA_miRgene_level: HDF5Matrix with 721 rows and 123 columns
 [3] RPPA_Gene_Level: HDF5Matrix with 166 rows and 123 columns
 [4] RNAseq_HiSeq_Gene_level: HDF5Matrix with 20190 rows and 123 columns
```

Columns:

-   Age \[int\]

-   Sex \[0/1\]

-   Response \[0/1\]

```{r}
# Load the MAE object from your .rds file
mae_tcga_kipan <- readRDS("~/tcga-kipan/tcga-kipan_mae_data/mae.rds")
mae <- mae_tcga_kipan

# # Check the class of the object
# class(mae)
# 
# # Print the summary of the MAE object
# summary(mae)
# 
# # List all assays (experiments) contained in the MAE object
# experiments(mae)
# 
# # Get a quick overview of the metadata associated with samples (colData)
# colData(mae)
# metadata(mae)

# Open the HDF5 file
h5_file <- HDF5Array::H5File("~/tcga-kipan/tcga-kipan_mae_data/experiments.h5")
h5_file

# Load the HDF5 file containing assay data (if it's separate from the RDS)
h5_contents <- h5ls(h5_file)

# get assay data from mrna dataset 
assay_data <- h5mread(h5_file, "assay004")

h5closeAll()
# Read the dataset with gene and sample dimnames
gene_dimnames_kipan <- h5read("~/tcga-kipan/tcga-kipan_mae_data/experiments.h5", "/.assay004_dimnames/1")
sample_dimnames<- h5read("~/tcga-kipan/tcga-kipan_mae_data/experiments.h5", "/.assay004_dimnames/2")

# convert assay data into a dataframe with genes as columns and sample names as rows
class(assay_data)
kipan_assay_df <- as.data.frame(t(assay_data))
rownames(kipan_assay_df) <- sample_dimnames
colnames(kipan_assay_df) <- gene_dimnames_kipan
```

## TCGA-THCA

Condition: Thyroid carcinoma

Number of Samples: 217

Experiments: 4 experiments

```         
[1] Methylation_Gene_level_HM450K: HDF5Matrix with 20118 rows and 217 columns
[2] miRNA_Gene_level: HDF5Matrix with 808 rows and 217 columns
[3] RPPA_Gene_Level: HDF5Matrix with 134 rows and 217 columns
[4] RNAseq_HiSeq_Gene_level: HDF5Matrix with 19927 rows and 217 columns
```

Columns:

-   Age \[int\]

-   Sex \[0/1\]

-   Response \[0/1\]

```{r}
# Load the MAE object from your .rds file
mae_tcga_thca <- readRDS("~/tcga-thca/tcga-thca_mae_data/mae.rds")
mae <- mae_tcga_thca

# # Check the class of the object
# class(mae)
# 
# # Print the summary of the MAE object
# summary(mae)
# 
# # List all assays (experiments) contained in the MAE object
# experiments(mae)
# 
# # Get a quick overview of the metadata associated with samples (colData)
# colData(mae)
# metadata(mae)

# Open the HDF5 file
h5_file <- HDF5Array::H5File("~/tcga-thca/tcga-thca_mae_data/experiments.h5")
h5_file

# Load the HDF5 file containing assay data (if it's separate from the RDS)
h5_contents <- h5ls(h5_file)

# get assay data from mrna dataset 
assay_data <- h5mread(h5_file, "assay004")

h5closeAll()
# Read the dataset with gene and sample dimnames
gene_dimnames_thca <- h5read("~/tcga-thca/tcga-thca_mae_data/experiments.h5", "/.assay004_dimnames/1")
sample_dimnames<- h5read("~/tcga-thca/tcga-thca_mae_data/experiments.h5", "/.assay004_dimnames/2")

# convert assay data into a dataframe with genes as columns and sample names as rows
class(assay_data)
thca_assay_df <- as.data.frame(t(assay_data))
rownames(thca_assay_df) <- sample_dimnames
colnames(thca_assay_df) <- gene_dimnames_thca
```

### HDF5Matrix

```         
Methylation_Gene_level_HM450K: 20118 217
miRNA_Gene_level             : 808 217
RPPA_Gene_Level              : 134 217
RNAseq_HiSeq_Gene_level      : 19927 217
```

## Loading DE_Prior and applying to Assay Data

```{r}

# loading the de_prior list
de_prior <- readr::read_delim("~/DE_Prior.txt", delim = "\t")

summary(de_prior)

hist(de_prior$DE_Prior_Rank)
```

### TCGA BLCA Dataset

```{r}

# checking the number of genes shared between de_prior list and assay data
common_genes <- de_prior[de_prior$Gene_Name %in% gene_dimnames_blca, ]
# head(common_genes)
summary(common_genes)

# checking the number of genes only in de_prior list and not assay data
genes_only_in_prior <- setdiff(de_prior$Gene_Name, gene_dimnames_blca)
# head(genes_only_in_prior)
# length(genes_only_in_prior)

# checking the number of genes only in assay data and not de_prior list
## this is the number of genes after the feature selection?
genes_only_in_assay <- setdiff(gene_dimnames_blca, de_prior$Gene_Name)
# head(genes_only_in_assay)
# length(genes_only_in_assay)

# creating a table with the number of common genes & number of genes in assay data after removing de_prior list
genes_info_table <- data.frame(dataset = "TCGA BLCA", common_genes_count = nrow(common_genes), assay_only_genes_count = length(genes_only_in_assay))

```

### TCGA BRCA Dataset

```{r}

# checking the number of genes shared between de_prior list and assay data
common_genes <- de_prior[de_prior$Gene_Name %in% gene_dimnames_brca, ]
head(common_genes)
summary(common_genes)

# checking the number of genes only in de_prior list and not assay data
genes_only_in_prior <- setdiff(de_prior$Gene_Name, gene_dimnames_brca)
head(genes_only_in_prior)
summary(genes_only_in_prior)

# checking the number of genes only in assay data and not de_prior list
## this is the number of genes after the feature selection?
genes_only_in_assay <- setdiff(gene_dimnames_brca, de_prior$Gene_Name)
head(genes_only_in_assay)
summary(genes_only_in_assay)

# adding to gene_info_table with the number of common genes & number of genes in assay data after removing de_prior list
genes_info_table <- rbind(genes_info_table, data.frame(dataset = "TCGA BRCA", common_genes_count = nrow(common_genes), assay_only_genes_count = length(genes_only_in_assay)))
```

### TCGA KIPAN Dataset

```{r}

# checking the number of genes shared between de_prior list and assay data
common_genes <- de_prior[de_prior$Gene_Name %in% gene_dimnames_kipan, ]
head(common_genes)
summary(common_genes)

# checking the number of genes only in de_prior list and not assay data
genes_only_in_prior <- setdiff(de_prior$Gene_Name, gene_dimnames_kipan)
head(genes_only_in_prior)
summary(genes_only_in_prior)

# checking the number of genes only in assay data and not de_prior list
## this is the number of genes after the feature selection?
genes_only_in_assay <- setdiff(gene_dimnames_blca, de_prior$Gene_Name)
head(genes_only_in_assay)
summary(genes_only_in_assay)

# adding to gene_info_table with the number of common genes & number of genes in assay data after removing de_prior list
genes_info_table <- rbind(genes_info_table, data.frame(dataset = "TCGA KIPAN", common_genes_count = nrow(common_genes), assay_only_genes_count = length(genes_only_in_assay)))
```

### TCGA THCA Dataset

```{r}

# checking the number of genes shared between de_prior list and assay data
common_genes <- de_prior[de_prior$Gene_Name %in% gene_dimnames_thca, ]
head(common_genes)
summary(common_genes)

# checking the number of genes only in de_prior list and not assay data
genes_only_in_prior <- setdiff(de_prior$Gene_Name, gene_dimnames_thca)
head(genes_only_in_prior)
summary(genes_only_in_prior)

# checking the number of genes only in assay data and not de_prior list
## this is the number of genes after the feature selection?
genes_only_in_assay <- setdiff(gene_dimnames_thca, de_prior$Gene_Name)
head(genes_only_in_assay)
summary(genes_only_in_assay)

# adding to gene_info_table with the number of common genes & number of genes in assay data after removing de_prior list
genes_info_table <- rbind(genes_info_table, data.frame(dataset = "TCGA THCA", common_genes_count = nrow(common_genes), assay_only_genes_count = length(genes_only_in_assay)))

print(genes_info_table)
```

### GSE71669 Dataset

```{r}
gene_dimnames_GSE71669 <- readr::read_delim("~/mapping_output.txt", delim = "\t")
# checking the number of genes shared between de_prior list and assay data
common_genes <- de_prior[de_prior$Gene_Name %in% gene_dimnames_GSE71669$hgnc_symbol, ]
# head(common_genes)
# summary(common_genes)

# checking the number of genes only in de_prior list and not assay data
genes_only_in_prior <- setdiff(de_prior$Gene_Name, gene_dimnames_GSE71669$hgnc_symbol)
# head(genes_only_in_prior)
# summary(genes_only_in_prior)

# checking the number of genes only in assay data and not de_prior list
## this is the number of genes after the feature selection?
genes_only_in_assay <- setdiff(gene_dimnames_GSE71669$hgnc_symbol, de_prior$Gene_Name)
# head(genes_only_in_assay)
# summary(genes_only_in_assay)

# adding to gene_info_table with the number of common genes & number of genes in assay data after removing de_prior list
genes_info_table <- rbind(genes_info_table, data.frame(dataset = "GSE71669", common_genes_count = nrow(common_genes), assay_only_genes_count = length(genes_only_in_assay)))

print(genes_info_table)
```
